---
output: html_document
editor_options: 
  chunk_output_type: console
---
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory

# Load your packages
library(tidyverse);library(lubridate);library(here);library(sf);library(ggnewscale)
here()

# Set your ggplot theme

# Load your datasets

#Air Quality
TXPM2523 <- read.csv(here("Data/TXPM2523.csv"))
TXPM2522 <- read.csv(here("Data/TXPM2522.csv"))
TXPM2521 <- read.csv(here("Data/TXPM2521.csv"))
TXPM2520 <- read.csv(here("Data/TXPM2520.csv"))
TXPM2519 <- read.csv(here("Data/TXPM2519.csv"))
TXOzone23 <- read.csv(here("Data/TXOzone23.csv"))
TXOzone22 <- read.csv(here("Data/TXOzone22.csv"))
TXOzone21 <- read.csv(here("Data/TXOzone21.csv"))
TXOzone20 <- read.csv(here("Data/TXOzone20.csv"))
TXOzone19 <- read.csv(here("Data/TXOzone19.csv"))

#eGRID
eGRID23_PLNT <- read.csv(here("Data/egrid2023_data_rev2_PLNT.csv"),
                         na.strings = c("", " ", "0")) %>% slice(-1) 

eGRID22_PLNT <- read.csv(here("Data/egrid2022_data_PLNT.csv"),
                         na.strings = c("", " ", "0")) %>% slice(-1) 

eGRID21_PLNT <- read.csv(here("Data/eGRID2021_data_PLNT.csv"),
                         na.strings = c("", " ", "0")) %>% slice(-1) 

eGRID20_PLNT <- read.csv(here("Data/eGRID2020_Data_v2_PLNT.csv"),
                         na.strings = c("", " ", "0")) %>% slice(-1) 

eGRID19_PLNT <- read.csv(here("Data/egrid2019_data_PLNT.csv"),
                         na.strings = c("", " ", "0")) %>% slice(-1) 

#ADD PRECIPITATION
```


# Rationale and Research Questions
This report examines how local conditions affect local air quality, focusing on power plant emissions and precipitation levels. While pollution is often studied at a global level, examining its contribution to climate change and warming temperatures, its affect communities local to the emissions source is studied less frequently. This analysis considers other factors (precipitation) that may influence air quality and explores how air quality levels vary day-to-day. Texas was chosen as the study area for this analysis because of its number of power plants and counties. This report investigates the following three questions:

1) Is there a correlation between county level emission levels from power plants and county air quality in Texas?
2) Is there a trend in air quality values over time in Harris county, Texas?
3) Is there a relationship between precipitation and air quality in Harris county Texas?

\newpage

# Dataset Information

This report analyzes three sources of data: EPA Air Quality, Emissions & Generation Resource Integrated Database (eGRID), and *INSERT PRECIPITATION DATA*

eGRID data is released annually and contains information for all power plants across the United States. It contains information about generation, age of the plant, fuel type, and emissions, including CO2 Equivalent emissions. This report uses the eGRID 2018 - 2023 annual reports and focuses only on Texas power plants. 

*INSERT SOMETHING ABOUT PRECIPITATION DATA*

EPA Air Quality Data is collected at local monitors across the United States. It can be downloaded by pollutant type, region, and year. This report uses data for Texas between 2018 and 2023 for PM2.5 and Ozone. PM2.5 and Ozone are two criteria pollutants that can cause respiratory issues, especially for people with pre-existing health conditions, children, and seniors. The AQI based on PM2.5 and Ozone are recorded separately and range on a scale from 0-500, with higher values being more dangerous. Values 0-50 indicate "Good" AQ, 50 - 100 "Moderate", and increasingly poor from there. For more information about Air Quality Index and criteria pollutants, visit https://www.epa.gov/outdoor-air-quality-data/air-data-basic-information

\newpage

# Exploratory Analysis 
```{r}

```


\newpage

# Analysis

## Question 1: Is there a correlation between county level emission levels from power plants and county air quality?
###How often is emissions data reported?

Insert map of reporting vs not reporting plants

###How are air quality stations spatially located compared to emissions reporting power plants?
Insert map of plants and air quality stations

###Are annual county emissions correlated with annual county air quality?
Insert charts of AQ vs O3 and AQ vs PM25
Statistical correlation test


## Question 2: 

## Question 3: 




\newpage

# Summary and Conclusions

The analysis of correlation between air quality and emissions at the county level per year did not result in significant results. There is not a significant correlation between annual county emission from power plants and average annual county AQI score from Ozone or PM2.5. Additionally, no county on average saw any poor or unsafe Air Quality Index levels. There are two potential factors influencing this lack of correlation. One is that air quality and emissions cannot be studied at localized scale. Likely emissions disperse out of the county it is in based on local weather patterns. Conducting analysis of local air quality based on power plant emissions might be more accurate if wind patterns were taken into consideration to estimate where emissions may be concentrating. Another reason there may not be a correlation is because there are not enough data points in this analysis to see a trend. All plants do not report emissions and there are not air quality stations in all counties. There were only 5 counties with a PM2.5 monitor and at least one emission reporting plant and only 5 counties with an Ozone monitor and at least oen emission reporting plant. To perform this analysis more accurately, data should be looked at country wide and there should be better standards for ensuring all active power plants report their emissions.

\newpage

# References
<add references here if relevant, otherwise delete this section> 
